The Hot Hand Fallacy in Call of Duty Search and Destroy

### 2022-08-26 paper/paper.Rmd
thursday morning review

cod_win_series_prop <- cod_rounds |> 
  distinct(series_id, win_series, n_rounds) |> 
  count(n_rounds) |> 
  mutate(prop = n / sum(n)) |> 
  arrange(n_rounds)
```

@@

1.  One team eliminates all members of the other team prior to a bomb plant. (Eliminating team wins.)
2.  The offensive team eliminates all members of the defensive team after a bomb plant. (Offense wins.)
3.  The defense defuses the bomb after a bomb plant.[^4] (Defense wins.)
4.  The offense does not make a plant by the time the round timer ends. (Defense wins.)

[^4]: Often the defensive team will try to eliminate all team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.

@@

We note that SnD is not the only game mode played in competitive CoD.[^5] Teams play in a head-to-head, best-of-five format, where SnD is always played as the second and fifth game modes. The best-of-five matchup could also be called a "series", but since we analyze only the SnD games, we refer to the SnD games as series.

[^5]: Hardpoint has been played as the first and fourth maps in a matchup. The variant for the third map was Domination in the 2020 season, and Control in the 2021 and 2022 seasons.

@@

CoD has roughly gone through three eras of professional gaming: (1) Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World League (CWL), initiated in 2016; and (3) the 12-franchise CoD League (CDL), operating since 2020. The CDL has completed three year-long "seasons" as of August 2022.[^6]

[^6]: CoD is fairly unique compared to other esports in that it runs on an annual lifecycle, with releases coming in the late fall. A new game under the same brand---Call of Duty---is published every year by a rotating set of developers. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been Modern Warfare (2020), Cold War (2021) and Vanguard (2022).

The data set consists of all SnD matches played in major tournaments and qualifiers during the CDL era, totaling 7,792 rounds across 852 series.[^7] Data has been collected in spreadsheets by community member "IOUTurtle".[^8]

[^7]: 288 of the series occur in major tournaments, which are considered to be more competitive than the qualifiers since they are played in person (COVID-permitting), eliminating randomness due to online latency and ping. The qualifiers are played online.

[^8]: Raw data: <https://linktr.ee/CDLArchive>. Author: <https://twitter.com/IOUTurtle>. Our clean data can be found at <https://github.com/tonyelhabr/fps_round_games/blob/master/data/cod_rounds.csv>.

6-0 sweeps make up 4.7% of all series, while 6-5 series make up 23.7% of series. The observed offensive round win percentage across all rounds is $\bar{\tau}_O$ = 47.8%.[^9] Table \ref{tbl:o-win-prop-by-series-state} shows offensive round win percentages by series "state", i.e. the number of round wins by each team prior to an upcoming round. Offensive round win percentage is not quite constant, although never veers more than 10% from $\bar{\tau}_O$.

[^9]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 1. 47.2% in MW (2020) 2. 47.9% in Cold War (2021) 3. 48.1% in Vanguard (2022)

@@

\caption{Offensive round win percentage for the upcoming round, given both the offensive and defensive team's prior number of round wins. Numbers in parentheses indicate sample sizes.}

@@

0 & 47.8\%(852) & 46.6\%(408) & 43.1\%(216) & 43.5\%(115) & 43.3\%(67)  & 40.5\%(37)  \\
1 & 48.6\%(444) & 49.3\%(418) & 51.5\%(309) & 43.4\%(205) & 43.3\%(120) & 39.4\%(99)  \\
2 & 52.8\%(218) & 48.9\%(305) & 48.9\%(315) & 46.6\%(262) & 48.7\%(189) & 42.1\%(133) \\
3 & 54.5\%(123) & 46.0\%(200) & 49.6\%(250) & 45.6\%(248) & 44.4\%(214) & 44.8\%(174) \\
4 & 56.9\%(65)  & 54.5\%(145) & 47.2\%(193) & 44.7\%(228) & 55.2\%(221) & 50.5\%(208) \\
5 & 47.4\%(38)  & 49.4\%(83)  & 47.1\%(136) & 50.9\%(175) & 45.2\%(177) & 46.0\%(202) \\

@@

\end{table}
```
At tournaments, higher-seeded ("better") teams choose whether they want to start on offense in the SnD series, played as the second or fifth game mode of the best-of-five matchup. Their opponent is assigned to start on offense in the SnD slot not chosen.[^10] Outside of tournaments, teams all play each other twice throughout the season, constituting qualifiers. Each team gets to play the higher-seeded role once in their head-to-head matchups with a given team, despite their win-loss records. Overall, we can argue that potential bias due to stronger teams playing defense---or, offense, if offense happens to be generally advantageous for a given map---is very minimal.

[^10]: Anecdotally, the higher-seeded teams tend to choose to play defense in the first round of the SnD series played as the second game mode in the best-of-five matchup, although this choice is not consistent across teams, or even with the same team over time.

@@

The concept of momentum goes hand-in-hand with a discussion of the nature of series outcomes.[^11] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and the "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

[^11]: We often use "streaks" and momentum interchangeably, but as Steeger et al. ([-@steeger2021]) note, momentum implies dependence between events, whereas streaking does not.

@@

Other research has borrowed techniques from the field of quality control, such as identifying unlikely sequences of events with the Wald-Wolfowitz runs test (Peel and Clauset 2015, Steeger et al. 2021). Peel and Clauset found no evidence for unlikely sequences of scoring events in the NHL, College Football (CFB), and National Football League (NFL), but did in the NBA. As a check on their entropy approach to momentum identification, Steeger et al. found several NHL teams with sequence of wins in the 2018-2019 regular season that violated the Wald-Wolfowitz null hypothesis.

@@

1.  Both have economic aspects that can create clear advantages for one team in a round, given how prior rounds played out. CoD has no such equivalent, except for perhaps "score streaks", which infrequently occur.
2.  Teams play either offense or defense for many consecutive rounds. (The former has a 15-15-1 format and the latter uses a 12-12-1 format.) On the other hand, teams in CoD SnD rotate roles every round, analogous to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series for professional sports like the MLB, NBA, and NHL.[^12] While theoretically one might be able to account for inter-round win percentage relationships in any kind of format, such as a 5-5-1, the rotation of team sides every round is convenient for convincing ourselves that rounds could reasonably be modeled as i.i.d. Bernoulli trials.
3.  Both Counter-Strike and Valorant have overtime rules---team must win by two rounds---which can make end-of-series sequences difficult to model properly. CoD SnD does not have overtime rules.

[^12]: 1-1-1-1-1-1-1 is not used today in these leagues, but it was at least once in each.

To our knowledge, there is no existing public statistical research on the CoD SnD format, beyond descriptive analysis on social media. While intra-round trends may be more directly applicable to teams looking for an advantage on their competition, broader investigation of a concept like the hot-hand fallacy in a sport where it has not yet been invetigated---particularly one that is less subjected to factors that may be difficult to control for, e.g. weather---should be useful as a precedent or supplement for future researchers.

@@

In a best-of-$s$ format, assuming a constant round win probability $\phi$[^13] (where $0 \leq \phi \leq 1$) for one team[^14], the expected proportion of series ending in $r$ rounds ($r \leq s$) is given by Equation \ref{eq:series-length}.

[^13]: If $\phi > 0.5$, we can say that this is the stronger team.

[^14]: $\phi$ and other symbols are selected so as to reserve symbols like $p$ for usage in other contexts without causing confusion for the reader. Upper-case symbol, e.g. $\Phi$, are consistently used in this paper to represent proportions, while lower-case symbols corresponding to upper-case symbols, e.g. $\phi$, are used to represent Bernoulli trial success rates. "Hats", e.g. $\hat{\Phi}$, are used to convey expectations, while bare symbols symbols convey observational data.

```{=tex}
\begin{equation}\label{eq:m}
m = \frac{s + 1}{2}.
\end{equation}
```
```{=tex}
\begin{equation}\label{eq:series-length}
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m).
\end{equation}
```
For example, assuming $\phi = 0.5$, the probability of a series ending in nine rounds in CoD SnD, where $s=11$, is

@@

```{=tex}
\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum_{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}
\end{equation}
```
in which $\hat{\Phi}(r)$ and $\Phi(r)$ represent the expected and observed round win proportions respectively, and $R = [6, 7, 8, 9, 10, 11]$ for CoD SnD. The null hypothesis is that the test statistic is chi-squared distributed, where the number of degrees of freedom is $\vert R \vert -1 = 5$ for Cod SnD. We reject the null hypothesis if $\chi^2 > \chi^2_{1-\alpha,\vert R \vert -1}$, where $\chi^2_{1-\alpha,\vert R \vert -1}$ is the chi-square critical value with $\vert R \vert -1$ degree of freedom and siginficance level $\alpha$.

@@

```{=tex}
\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 = \begin{cases} 
  \frac{m}{r} & \text{team wins series}, \\ 
  \frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}
```
using $m$ from Equation \ref{eq:m}.

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than $p^{+|r}$.[^15]

[^15]: Unfortunately, there does not exist a closed form representation for the expected value of $\hat{P}^{+|k,r}_{MS}$ for $k > 1$. Nonetheless, one may run simulations to estimate the expected value, as we choose to do. We have adapted the code from Vafa [-@vafa2017], which implements MS's framework [-@miller2018].

Although one might be tempted to mirror the equal proportions hypothesis testing performed by MS, our context is fundamentally different from that of MS, who focus on longitudinal data in controlled settings, unaffected by opposition.[^16] The number of trials is fixed in their experimental designs, but in CoD SnD, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series. The Bernoulli trial success probability, i.e. single round win percentage in CoD SnD, is not independent of the opponent. Consequently, a statistical test of the difference in $\hat{P}^{+|k,r}$ and $\hat{P}^{-|kr}$, as performed by MS to evaluate their hypothesis regarding post-streak success rate, is not completely appropriate, although useful as a reference.

[^16]: GVT also perform statistical tests on shots from players in live games, i.e. "observational" data, but they note that their findings are likely affected player shot selection in the face of defensive strategy by the opposing team.

@@

```{=tex}
\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 = \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, \\
  \frac{s - m - k}{r - k} & \text{team loses series},
\end{cases}
\end{equation}
```
using $m$ from Equation \ref{eq:m}.

@@

```{=tex}
\begin{equation}\label{eq:binom}
\Pr(r^w) = {\binom {r}{r^+}} \omega^{r^+}(1-\omega)^{r-r^+}.
\end{equation}
```
Treating the notional proportion $\hat{P}^{+|k,r}_0$ as the null $\omega_0$ and plugging in the observed proportion $P^{+|k,r}$ for $\omega$ in Equation \ref{eq:binom} (treating proportions as probabilities), we can evaluate the null hypothesis that the observed proportion of post-streak round wins is equal to the notional proportion, given streak length $k$ and series length $r$. If we can reject this null hypothesis, then we can consider team momentum, represented by post-streak success, plausible.

@@

and where $N^\zeta$ and $\hat{N}^\zeta$ are the observed and expected number of sequences respectively. The null hypothesis is that $z$ is normally distributed. If $z > +z_{(1-\alpha)/2}$ or $z < -z_{(1-\alpha)/2}$ for critical values at a significance level $\alpha$, we can reject the null hypothesis for such sequences and build an argument in support of the hot hand effect.

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the new $\hat{\Phi}(r)$ when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$, resulting in a new set of expected proportions of series lasting $r$ rounds $\hat{\Phi}_2(r)$.[^17] We observe that $\hat{\Phi}_2(r)$ is larger than $\hat{\Phi}_0(r)$ for $r \in [6, 7]$, more closely matching $\Phi(r)$. $\hat{\Phi}_2(r)$ is also closer to the observed $\Phi(r)$ for $r \in [9, 10]$, although not for $r \in [8, 11]$.

[^17]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins.[^18] Table \ref{tbl:pwkr} compares the notional and MS expected proportions, $\hat{P}^{+|k,r}_0$ and $\hat{P}^{+|k,r}_{MS}$ respectively, with the observed round win proportion, $P^{+|kr}$, following streaks of $k=3$ round wins given the series outcome.

[^18]: Three happens to also be a reasonable number for series that last at maximum 11 rounds.

@@

\end{table}
```
With the exception of $\hat{P}^{+|k,r}_{MS}$ when $r = 10$ and the eventual series loser is the team that wins after a streak of three round wins, all un-adjusted binomial test $p$-values are greater than the $\alpha = 0.05$ confidence level, implying that we cannot reject the null hypothesis that the expected notional and MS post-streak round win proportions are different than the observed proportion. However, after adjusting the $p$-values with the Benjamini and Yekutieli (BY) correction [-@benjamini2001], the null hypothesis cannot be rejected for any case.[^19]

[^19]: As noted before, the MS expectations may be unreliable for CoD SnD, so one is inclined to prefer the results of the notional binomial tests.

When performing the same tests for streaks of two, four, and five, there is no case in which we can reject the binomial null hypothesis for the expected notional proportion $\hat{P}^{+|k,r}_0$ (even before the BY p-value correction). The null hypothesis can only be rejected for the expected MS proportion $\hat{P}^{+|k,r}_{MS}$ when $k = 5, r = 7$ after applying the BY p-value correction.[^20]

[^20]: At 36, the sample size for $k = 5, r = 7$ is not as large as most other combinations of $k$ and $r$.

@@

Now let us explicitly consider the round, $i$. As shown in Table \ref{tbl:pw3ri}, the notional win proportion in the round $i$ immediately following a streak of $k$ round wins in a series lasting $r$ rounds, $\hat{P}{+|k,r,i}_0$, is statistically different than the observed $P^{+|k,r,i}$ in several cases, notably when $r = i$ for each of $r \in [9, 10, 11]$.[^21] Interestingly, the observed rate is greater than the expected rate in each case where we can reject the null hypothesis, whereas the observed rate is less than the expected rate in all other cases, as well in a strong majority of cases not shown.

[^21]: The null can also be rejected for $r \in [7, 8]$, although these are not shown, and the sample sizes are smaller.

@@

\end{table}
```
```{r}
#| label: cod_round_arrangements

simulate_cod_round_arrangements <- function(p = 0.5, n_sims = 10000, seed = 42) {
  withr::local_seed(seed)
  w <- sample(c(0, 1), size = MAX_COD_ROUND * n_sims, replace = TRUE, prob = c(1 - p, p))
  m <- matrix(w, nrow = n_sims, ncol = n_rounds_max)
  # data.frame(x)
  df <- as_tibble(m)
  names(df) <- sprintf('%d', 1:n_rounds_max)
  df$i <- 1:nrow(df)

  sim_rounds <- tibble(
    i = rep(1:n_sims, each = n_rounds_max),
    r = rep(1:n_sims, times = n_rounds_max),
    w = w
  ) |> 
    group_by(i) |> 
    mutate(
      cumu_w = cumsum(w),
      cumu_l = cumsum(w == 0)
    ) |> 
    ungroup() |> 
    mutate(
      cumu_wl_max = ifelse(cumu_w > cumu_l, cumu_w, cumu_l)
    ) |> 
    filter(cumu_wl_max <= 6)

  sim_rounds <- df |> 
    pivot_longer(
      -i,
      names_to = 'r',
      values_to = 'w'
    ) |> 
    mutate(
      across(r, as.integer)
    ) |> 
    group_by(i) |> 
    mutate(
      cumu_w = cumsum(w),
      cumu_l = cumsum(w == 0)
    ) |> 
    ungroup() |> 
    mutate(
      cumu_wl_max = ifelse(cumu_w > cumu_l, cumu_w, cumu_l)
    ) |> 
    filter(cumu_wl_max <= 6)

  sim_rounds_to_drop <- anti_join(
    sim_rounds |> 
      filter(cumu_wl_max == 6L),
    sim_rounds |> 
      filter(cumu_wl_max == 6L) |> 
      group_by(i) |> 
      slice_min(r, n = 1) |> 
      ungroup(), 
    by = c('i', 'r', 'w', 'cumu_w', 'cumu_l', 'cumu_wl_max')
  )

@@

  ## always from offensive perspective
  pre_expected_records <- sim_rounds |> 
    inner_join(
      sim_rounds |> 
        filter(cumu_wl_max == 6) |> 
        mutate(
          zeros_win = cumu_l == cumu_wl_max
        ) |> 
        select(i, zeros_win),
      by = 'i'
    ) |> 
    mutate(
      across(w, ~ifelse(zeros_win, abs(1 - .x), .x))
    )

>>>>

cod_win_series_prop <- cod_rounds |> 
  distinct(series_id, win_series, n_rounds) |> 
  count(n_rounds) |> 
  mutate(prop = n / sum(n)) |> 
  arrange(n_rounds)

cod_rounds |> 
  filter(round == 1) |> 
  count(game, team, is_offense) |> 
  group_by(game, team) |> 
  mutate(prop = n / sum(n)) |> 
  ungroup() |>
  filter(is_offense) %>% 
  arrange(-prop)
```

@@

1.  One team eliminates all members of the other team prior to a bomb plant. (Eliminating team wins.)
2.  The offensive team eliminates all members of the defensive team after a bomb plant. (Offense wins.)
3.  The defense defuses the bomb after a bomb plant. (Defense wins.)
4.  The offense does not make a plant by the time the round timer ends. (Defense wins.)

@@

We note that SnD is not the only game mode played in competitive CoD.[^4] Teams play in a head-to-head, best-of-five format, where SnD is always played as the second and fifth game modes. The best-of-five matchup could also be called a "series", but since we analyze only the SnD games, we refer to the SnD games as series.

[^4]: Hardpoint has been played as the first and fourth game modes in a matchup. The third game mode was Domination in the 2020 season, and Control in the 2021 and 2022 seasons.

@@

The data set consists of all SnD matches played in major tournaments and qualifiers during the past three years[^5], totaling 7,792 rounds across 852 series.[^6] Data has been collected in spreadsheets by community member "IOUTurtle".[^7]

[^5]: CoD has roughly gone through three eras of professional gaming: (1) Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World League (CWL), initiated in 2016; and (3) the 12-franchise CoD League (CDL), operating since 2020. The CDL has completed three year-long "seasons" as of August 2022.

    CoD is fairly unique compared to other esports in that it runs on an annual lifecycle, with releases coming in the late fall. A new game under the same brand---Call of Duty---is published every year by a rotating set of developers. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been Modern Warfare (2020), Cold War (2021) and Vanguard (2022).

[^6]: 288 of the series occur in major tournaments, which are considered to be more competitive than the qualifiers since they are played in person (COVID-permitting), eliminating randomness due to online latency and ping. The qualifiers are played online.

[^7]: Raw data: <https://linktr.ee/CDLArchive>. Cleaned data: <https://github.com/%7Bauthor%7D/%7Brepo%7D/blob/master/data/%7Bfile%7D.csv>.

Sweeps (6-0 series) make up 4.7% of all series, while 6-5 series make up 23.7% of series. The observed offensive round win percentage across all rounds is $\bar{\tau}$ = 47.8%.[^8] Table \ref{tbl:o-win-prop-by-series-state} shows offensive round win percentages by series "state", i.e. the number of round wins by each team prior to an upcoming round. Offensive round win percentage is not quite constant, although never veers more than 10% from $\bar{\tau}$.

[^8]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 1. 47.2% in MW (2020) 2. 47.9% in Cold War (2021) 3. 48.1% in Vanguard (2022)

@@

\caption{Offensive round win percentage (\%) for the upcoming round, given both the offensive and defensive team's prior number of round wins. Numbers in parentheses indicate sample sizes.}

@@

0 & 47.8 (852) & 46.6 (408) & 43.1 (216) & 43.5 (115) & 43.3 (67)  & 40.5 (37)  \\
1 & 48.6 (444) & 49.3 (418) & 51.5 (309) & 43.4 (205) & 43.3 (120) & 39.4 (99)  \\
2 & 52.8 (218) & 48.9 (305) & 48.9 (315) & 46.6 (262) & 48.7 (189) & 42.1 (133) \\
3 & 54.5 (123) & 46.0 (200) & 49.6 (250) & 45.6 (248) & 44.4 (214) & 44.8 (174) \\
4 & 56.9 (65)  & 54.5 (145) & 47.2 (193) & 44.7 (228) & 55.2 (221) & 50.5 (208) \\
5 & 47.4 (38)  & 49.4 (83)  & 47.1 (136) & 50.9 (175) & 45.2 (177) & 46.0 (202) \\

@@

\end{table}
```

The professional competition format balances the frequency with which teams play offense or defense to start a series, to the extent possible. Thus, win percentages in certain states are subjected to minimal selection bias, i.e. bias due to better teams playing defense more frequently.[^9]

[^9]: At tournaments, higher-seeded ("better") teams choose whether they want to start on offense in the SnD series, played as the second or fifth game mode of the best-of-five matchup. Their opponent is assigned to start on offense in the SnD slot not chosen. Anecdotally, the better team tends to choose to play defense in the first round of the SnD series played as the second game mode in the best-of-five matchup, although this choice is not consistent across teams, or even with the same team over time.

    Outside of tournaments, teams all play each other twice throughout the season, constituting qualifiers. Each team gets to play the higher-seeded role once in their head-to-head matchups with a given team, despite their win-loss records.

@@

The concept of momentum goes hand-in-hand with a discussion of the nature of series outcomes.[^10] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and the "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

[^10]: We often use "streaks" and momentum interchangeably, but as Steeger et al. [-@steeger2021] note, momentum implies dependence between events, whereas streaking does not.

@@

Other research has borrowed techniques from the field of quality control, such as identifying unlikely sequences of events with the Wald-Wolfowitz runs test (Peel and Clauset 2015, Steeger et al. 2021). Peel and Clauset found no evidence for unlikely sequences of scoring events in the NHL, College Football, and National Football League, although did in the NBA. As a check on their entropy approach to momentum identification, Steeger et al. found several NHL teams with sequences of wins in the 2018-2019 regular season that violated the Wald-Wolfowitz null hypothesis.

@@

1.  Both have economic aspects that can create clear advantages for one team in a round, given how prior rounds played out. CoD has no such equivalent, except for perhaps "score streaks", which infrequently occur.
2.  Teams play either offense or defense for many consecutive rounds. (The former has a 15-15-1 format and the latter uses a 12-12-1 format.) On the other hand, teams in CoD SnD rotate roles every round, analogous to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series for professional sports like the MLB, NBA, and NHL.[^11] While theoretically one might be able to account for inter-round win percentage relationships in any kind of format, such as a 5-5-1, the rotation of team sides every round is convenient for convincing ourselves that rounds could reasonably be modeled as i.i.d. Bernoulli trials.
3.  Both Counter-Strike and Valorant have overtime rules---a team must win by two rounds---which can make end-of-series sequences difficult to model properly. CoD SnD does not have overtime rules.

[^11]: 1-1-1-1-1-1-1 is not used today in these leagues, but it was at least once in each.

To our knowledge, there is no existing public statistical research on the CoD SnD format, beyond descriptive analysis on social media. While intra-round trends may be more directly applicable to teams looking for an advantage on their competition, broader investigation of a concept like the hot-hand fallacy in a sport where it has not yet been invetigated---particularly one that is less subjected to factors that may be difficult to control for, e.g. weather---should be useful as an inspiration for future researchers.

@@

In a best-of-$s$ format, assuming a constant round win probability $\phi$ (where $0 \leq \phi \leq 1$)[^12] for one team[^13], the expected proportion of series ending in $r$ rounds ($r \leq s$) is given by Equation \ref{eq:series-length}.

[^12]: If $\phi > 0.5$, we can say that this is the stronger team.

[^13]: $\phi$ and other symbols are selected so as to reserve symbols like $p$ for usage in other contexts without causing confusion for the reader. Upper-case symbol, e.g. $\Phi$, are consistently used in this paper to represent proportions, while lower-case symbols corresponding to upper-case symbols, e.g. $\phi$, are used to represent Bernoulli trial success rates. "Hats", e.g. $\hat{\Phi}$, are used to convey expectations, while bare symbols symbols convey observational data.

```{=tex}
\begin{equation}\label{eq:m}
m = \frac{s + 1}{2}.
\end{equation}
```
```{=tex}
\begin{equation}\label{eq:series-length}
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m).
\end{equation}
```
For example, assuming $\phi = 0.5$, the probability of a series ending in exactly nine rounds in CoD SnD, where $s=11$ (implying $m=6$), is

@@

```{=tex}
\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum_{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}
\end{equation}
```
in which $\hat{\Phi}(r)$ and $\Phi(r)$ represent the expected and observed round win proportions respectively, and $R = [6, 7, 8, 9, 10, 11]$ for CoD SnD. The null hypothesis is that the test statistic is chi-squared distributed, where the number of degrees of freedom is $\vert R \vert -1 = 5$ for Cod SnD.

@@

```{=tex}
\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 := \begin{cases} 
  \frac{m}{r} & \text{team wins series}, \\ 
  \frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}
```
for $r \in R$ and $m$ from Equation \ref{eq:m}.

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than $p^{+|r}$. Although one might be tempted to mirror the equal proportions hypothesis testing performed by MS, our context is fundamentally different from that of MS, who focus on longitudinal data in controlled settings, unaffected by opposition.[^14] The number of trials is fixed in their experimental designs, but in CoD SnD, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series. The Bernoulli trial success probability, i.e. single round win percentage in CoD SnD, is not independent of the opponent. Consequently, a statistical test of the difference in $\hat{P}^{+|k,r}$ and $\hat{P}^{-|kr}$, as performed by MS to evaluate their hypothesis regarding post-streak success rate, is not completely appropriate, although useful as a reference.

[^14]: GVT also perform statistical tests on shots from players in live games, i.e. "observational" data, but they note that their findings are likely affected by player shot selection in the face of defensive strategy by the opposing team.

@@

```{=tex}
\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 := \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, \\
  \frac{s - m - k}{r - k} & \text{team loses series},
\end{cases}
\end{equation}
```
for $r \in R$ and $m$ from Equation \ref{eq:m}.

@@

```{=tex}
\begin{equation}\label{eq:binom}
\binom {r}{r^+} \omega^{r^+}(1-\omega)^{r-r^+}.
\end{equation}
```
Treating the notional proportion $\hat{P}^{+|k,r}_0$ as the null $\omega_0$ and plugging in the observed proportion $P^{+|k,r}$ for $\omega$ in Equation \ref{eq:binom} (treating proportions as probabilities), we can evaluate the null hypothesis that the observed proportion of post-streak round wins is equal to the notional proportion, given streak length $k$ and series length $r$. If we can reject this null hypothesis, then we can consider team momentum, represented by post-streak success, plausible.

@@

and where $N^\zeta$ and $\hat{N}^\zeta$ are the observed and expected number of sequences respectively. The null hypothesis is that $z$ is normally distributed. If $z > +z_{(1-\alpha)/2}$ or $z < -z_{(1-\alpha)/2}$ at a significance level $\alpha$, we can reject the null hypothesis for such sequences and build an argument in support of the hot hand effect.

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the new $\hat{\Phi}(r)$ when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$, resulting in a new set of expected proportions of series lasting $r$ rounds $\hat{\Phi}_2(r)$.[^15] We observe that $\hat{\Phi}_2(r)$ is larger than $\hat{\Phi}_0(r)$ for $r \in [6, 7]$, more closely matching $\Phi(r)$. $\hat{\Phi}_2(r)$ is also closer to the observed $\Phi(r)$ for $r \in [9, 10]$, although not for $r \in [8, 11]$.

[^15]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins.[^16] Table \ref{tbl:pwkr} compares the notional and MS expected proportions, $\hat{P}^{+|k,r}_0$ and $\hat{P}^{+|k,r}_{MS}$ respectively, with the observed round win proportion, $P^{+|kr}$, following streaks of $k=3$ round wins given the series outcome.

[^16]: Three happens to also be a reasonable number for series that last at maximum 11 rounds.

@@

\end{table}
```
With the exception of $\hat{P}^{+|k,r}_{MS}$ when $r = 10$ and the eventual series loser is the team that wins after a streak of three round wins, all un-adjusted binomial test $p$-values are greater than the $\alpha = 0.05$ confidence level, implying that we cannot reject the null hypothesis that the expected notional and MS post-streak round win proportions are different than the observed proportion. However, after adjusting the $p$-values with the Benjamini and Yekutieli (BY) correction [-@benjamini2001], the null hypothesis cannot be rejected for any case.[^17]

[^17]: As noted before, the MS expectations may be unreliable for CoD SnD, so one is inclined to prefer the results of the notional binomial tests.

When performing the same tests for streaks of two, four, and five, there is no case in which we can reject the binomial null hypothesis for the expected notional proportion $\hat{P}^{+|k,r}_0$ (even before the BY p-value correction). The null hypothesis can only be rejected for the expected MS proportion $\hat{P}^{+|k,r}_{MS}$ when $k = 5, r = 7$ after applying the BY p-value correction.[^18]

[^18]: At 36, the sample size for $k = 5, r = 7$ is not as large as most other combinations of $k$ and $r$.

@@

Now let us explicitly consider the round, $i$. As shown in Table \ref{tbl:pw3ri}, the notional win proportion in the round $i$ immediately following a streak of $k$ round wins in a series lasting $r$ rounds, $\hat{P}{+|k,r,i}_0$, is statistically different than the observed $P^{+|k,r,i}$ in several cases, notably when $r = i$ for each of $r \in [9, 10, 11]$.[^19] Interestingly, the observed rate is greater than the expected rate in each case where we can reject the null hypothesis, whereas the observed rate is less than the expected rate in all other cases, as well in a strong majority of cases not shown.

[^19]: The null can also be rejected for $r \in [7, 8]$, although these are not shown, and the sample sizes are smaller.

@@

\end{table}
```
```{r}
#| label: cod_round_arrangements
simulate_cod_round_arrangements <- function(p = 0.5, n_sims = 10000, seed = 42) {
  withr::local_seed(seed)
  w <- sample(c(0, 1), size = MAX_COD_ROUND * n_sims, replace = TRUE, prob = c(1 - p, p))
  m <- matrix(w, nrow = n_sims, ncol = MAX_COD_ROUND)
  df <- as_tibble(m)
  names(df) <- sprintf('%d', 1:MAX_COD_ROUND)
  df$i <- 1:nrow(df)

  sim_rounds <- tibble(
    i = rep(1:n_sims, each = MAX_COD_ROUND),
    r = rep(1:n_sims, times = MAX_COD_ROUND),
    w = w
  ) |> 
    group_by(i) |> 
    mutate(
      cumu_w = cumsum(w),
      cumu_l = cumsum(w == 0)
    ) |> 
    ungroup() |> 
    mutate(
      cumu_wl_max = ifelse(cumu_w > cumu_l, cumu_w, cumu_l)
    ) |> 
    filter(cumu_wl_max <= COD_WIN_THRESHOLD)

  sim_rounds <- df |> 
    pivot_longer(
      -i,
      names_to = 'r',
      values_to = 'w'
    ) |> 
    mutate(
      across(r, as.integer)
    ) |> 
    group_by(i) |> 
    mutate(
      cumu_w = cumsum(w),
      cumu_l = cumsum(w == 0)
    ) |> 
    ungroup() |> 
    mutate(
      cumu_wl_max = ifelse(cumu_w > cumu_l, cumu_w, cumu_l)
    ) |> 
    filter(cumu_wl_max <= COD_WIN_THRESHOLD)

  sim_rounds_to_drop <- anti_join(
    sim_rounds |> 
      filter(cumu_wl_max == COD_WIN_THRESHOLD),
    sim_rounds |> 
      filter(cumu_wl_max == COD_WIN_THRESHOLD) |> 
      group_by(i) |> 
      slice_min(r, n = 1) |> 
      ungroup(), 
    by = c('i', 'r', 'w', 'cumu_w', 'cumu_l', 'cumu_wl_max')
  )

@@

  ## always from offensive perspective
  pre_expected_records <- sim_rounds |> 
    inner_join(
      sim_rounds |> 
        filter(cumu_wl_max == COD_WIN_THRESHOLD) |> 
        mutate(
          zeros_win = cumu_l == cumu_wl_max
        ) |> 
        select(i, zeros_win),
      by = 'i'
    ) |> 
    mutate(
      across(w, ~ifelse(zeros_win, abs(1 - .x), .x))
      # across(w, ~ifelse(zeros_win, '-', '+'))
    )
